---
title: 'MUSA500 Homework 3: The Application of Logistic Regression to Examine the Predictors of Car Crashes Caused by Alcohol'
author: "Ling Chen, Hang Zhao, Jiahang Li"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    code_folding: hide
    number_sections: yes
    code_download: no
    theme: united
    highlight: espresso
  pdf_document:
    toc: yes
editor_options:
  markdown:
    wrap: 72
---

## Set Up

We install a number of packages that we need in order to run logistic regression in R and specify the directory where our data are stored.

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "D:/00Penn-学习/MUSA500/HW3")
options(scipen=999)
```

```{r warning=FALSE, message=FALSE, cache=FALSE}
#install.packages("aod")
#install.packages("ggplot2")
#install.packages("rms")
#install.packages("gmodels")
#install.packages("nnet")
#install.packages("DAAG")
#install.packages("ROCR")
#install.packages("xtable")

library(aod)
library(ggplot2)
library(corrplot)
library(rms)
library(gmodels)
library(nnet)
library(DAAG)
library(ROCR)
library(xtable)
library(knitr)
library(kableExtra)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(purrr)
library(tidyr)

palette2 <- c("#bae4bc","#0868ac")

```

## Introduction

Our car crash data set has 43,364 observations. The relevant variables are described below:

-- Dependent Variable: ` DRINKING_D`, Drinking driver indicator (1 = Yes, 0 = No)
-- Predictor: ` FATAL_OT_M`, Crash resulted in fatality or major injury (1 = Yes, 0 = No)
-- Predictor: ` OVERTURNED`, Crash involved an overturned vehicle (1 = Yes, 0 = No)
-- Predictor: ` CELL_PHONE`, Driver was using cell phone (1= Yes, 0 = No)
-- Predictor: ` SPEEDING`, Crash involved speeding car (1 = Yes, 0 = No)
-- Predictor: ` AGGRESSIVE`, Crash involved aggressive driving (1 = Yes, 0 = No)
-- Predictor: ` DRIVER1617`, Crash involved at least one driver who was 16 or 17 years old (1 = Yes, 0 = No)
-- Predictor: ` DRIVER65PLUS`, Crash involved at least one driver who was at least 65 years old (1 = Yes, 0 = No)
-- Predictor: ` PCTBACHMOR`,% of individuals 25 years of age or older who have at least a bachelor’s degree in the Census Block Group where the crash took place
-- Predictor: ` MEDHHINC`, Median household income in the Census Block Group where the crash took place


```{r warning=FALSE, message=FALSE, cache=FALSE}
mydata <- read.csv("D:/00Penn-学习/MUSA500/HW3/Logistic Regression Data.csv")
head(mydata)
```

## Tabulation of the Dependent Variable

Let's look at the tabulation of our binary dependent variable, ` DRINKING_D`.

```{r warning=FALSE, message=FALSE, cache=FALSE}
summary(mydata)
table(mydata$DRINKING_D)
prop.table(table(mydata$DRINKING_D))
```

We see that there are 94.3% drivers who drink while driving, and only 5.7% who don't drink. 
The probability of drivers who drink while driving can be calculated using the formula

$$Probability(DRINKING) = \frac{Number \; of \; Drinking \; Driver}{Total \; Number \; of \; Drivers} = \frac{40879}{43364} = .94. $$

Similarly, the odds of drinking drivers can be calculated using the formula

$$Odds(DRINKING) = \frac{Number \; of \; Drinking \; Drivers}{Number \; of \; Non-drinking \; Drivers} = \frac{40879}{2485} = 16.45. $$
```{r categorical, message=FALSE, warning=FALSE, cache=FALSE}
FATAL_Chisq <- CrossTable(mydata$DRINKING_D, mydata$FATAL_OR_M, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE,chisq = TRUE)

OVER_Chisq <-CrossTable(mydata$DRINKING_D, mydata$OVERTURNED, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE,chisq = TRUE)

CELL_Chisq <- CrossTable(mydata$DRINKING_D, mydata$CELL_PHONE, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE,chisq = TRUE)

SPEED_Chisq <- CrossTable(mydata$DRINKING_D, mydata$SPEEDING, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE,chisq = TRUE)

AGGRE_Chisq <- CrossTable(mydata$DRINKING_D, mydata$AGGRESSIVE, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE,chisq = TRUE)

DRIVER1617_Chisq <- CrossTable(mydata$DRINKING_D, mydata$DRIVER1617, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE,chisq = TRUE)

DRIVER65_Chisq <-CrossTable(mydata$DRINKING_D, mydata$DRIVER65PLUS, prop.r=FALSE, prop.t=FALSE, prop.chisq=FALSE,chisq = TRUE)

chisq_values <- list(
  FATAL_Chisq$chisq,
  OVER_Chisq$chisq,
  CELL_Chisq$chisq,
  SPEED_Chisq$chisq,
  AGGRE_Chisq$chisq,
  DRIVER1617_Chisq$chisq,
  DRIVER65_Chisq$chisq
)
```

Prior to the predictive modeling, we use the Chi-Square test to determine whether the distribution of the categorical variable varies with respect to the drunk driving. From the table below, we can see that the majority of the variables except cell phone have a P-value lower than 0.05. Therefore, we can reject the null hypothesis and confirm that there are associations between drunk driving and overturned vehicle, speeding car, aggressive driving,young driver, old drivers, crash fatalities.
```{r binary_explore, echo=FALSE, warning=FALSE, cache=FALSE}

explore <- data.frame(
  Category = c("FATAL_OR_M", "OVERTURNED", "CELL_PHONE", "SPEEDING", "AGGRESSIVE", "DRIVER1617", "DRIVER65PLUS"),
  Drinking_N = c(188, 110, 28, 260, 916, 12, 119), 
  Drinking_Perc = c(NA, NA, NA, NA, NA, NA, NA), 
  Non_Drinking_N = c(1181, 612, 426, 1261, 18522, 674, 4237), 
  Non_Drinking_Perc = c(NA, NA, NA, NA, NA, NA, NA), 
  Total_N = c(1369, 722, 454, 1521, 19438, 686, 4356) ,
  p_value = c(0.00000000000000000000000000000000000002522202 ,0.0000000000000000000000000001551762 , 0.6872569 , 0.000000000000000000000000000000000000000000000000000000000000000000000000000000000006249562 , 0.000000000000000200079 , 0.000006115619 , 0.000000000000000000275703) 
)

explore <- explore %>%
  mutate(Drinking_Perc = 100* Drinking_N/43364,
         Non_Drinking_Perc=100*Non_Drinking_N/43364)

kable(explore, caption = "Traffic Accident Statistics - Categorical Variables") %>% 
  kable_styling()

```

```{r continuous_explore, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}

PCTBACHMOR_mean <- tapply(mydata$PCTBACHMOR,mydata$DRINKING_D,mean)
PCTBACHMOR_sd <- tapply(mydata$PCTBACHMOR,mydata$DRINKING_D,sd)
MEDHHINC_mean <- tapply(mydata$MEDHHINC,mydata$DRINKING_D,mean)
MEDHHINC_sd <- tapply(mydata$MEDHHINC,mydata$DRINKING_D,sd)

t.test(mydata$PCTBACHMOR~mydata$DRINKING_D)
t.test(mydata$MEDHHINC~mydata$DRINKING_D)

```

For the continuous variables `PCTBACHMOR`,`MEDHHINC`, we further examine whether their means differ for drunk driving or non-drunk driving. By using the independent samples t-test, the P-value for both variables are not statistically significantly different for crashes that involve drunk drivers and crashes that don't. Therefore, it can be concluded that the average values of the variables `PCTBACHMOR` and `MEDHHINC` are the same for accidents involving drunk drivers and those not involving them.

```{r continuous, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
explore_continuous <- data.frame(
  Category = c("PCTBACHMOR", "MEDHHINC"),
  Drinking_mean = c(16.61173, 31998.75 ),
  Drinking_sd = c(18.72091, 17810.5),
  Non_Drinking_mean = c(16.56986, 31483.05),
  Non_Drinking_sd = c(18.21426, 16930.1),
  P_value = c(0.9137,0.16)
)


kable(explore_continuous, caption = "Traffic Accident Statistics - Continuous Variables") %>% 
  kable_styling()
```

    
Among the independent variables, drunk drivers tend to have higher fatality rate, faster car speed, and higher percentage of overturning cars from the primary exploratory analysis.

```{r Pearson_Cor, echo=FALSE, fig.height=6, fig.width=8, message=FALSE, warning=FALSE, cache=FALSE}
mydata %>%
  dplyr::select(DRINKING_D, FATAL_OR_M, OVERTURNED, CELL_PHONE, SPEEDING, AGGRESSIVE, DRIVER1617, DRIVER65PLUS,PCTBACHMOR, MEDHHINC) %>%
   mutate(DRINKING_D = factor(DRINKING_D)) %>%
  gather(Variable, value, -DRINKING_D) %>%
    ggplot(aes(DRINKING_D, value, fill=DRINKING_D)) + 
      geom_bar(position = "dodge", stat = "summary", fun = "mean") + 
      facet_wrap(~Variable, scales = "free") +
      scale_fill_manual(values = palette2) +
      labs(x="Drinking/Non-drinking drivers", y="Value", 
           title = "Feature associations with drunk drivers") +
      theme(legend.position = "none")
```


The correlation matrix in the graph and table is shown below. It is used to test multicollinearity between the predictors. The table shows that the greatest correlation coefficients between predictors are 0.50 and 0.47, which are not strongly correlated. Therefore, no severe multicollinearity between the predictors has been observed.
```{r matrix,echo=FALSE, fig.height=4, fig.width=4}
# correlation matrix
data_cor <- mydata %>%
  dplyr::select(FATAL_OR_M, OVERTURNED, CELL_PHONE, SPEEDING, AGGRESSIVE, DRIVER1617, DRIVER65PLUS,PCTBACHMOR, MEDHHINC) 

cor_matrix <- cor(data_cor, method=c("pearson"))
corrplot(cor_matrix)
```

```{r table_matrix,echo=FALSE, fig.height=4, fig.width=4}
mydata_cor_dependent <- mydata %>%
dplyr::select(FATAL_OR_M, OVERTURNED, CELL_PHONE, SPEEDING, AGGRESSIVE, DRIVER1617, DRIVER65PLUS,PCTBACHMOR,MEDHHINC)

cor(mydata_cor_dependent, method=c("pearson"))

```

